---
title: "Ali's Amante Barista Stats"
header-includes:
   - \usepackage{multicol}
output:
  pdf_document:
    latex_engine: pdflatex
---
## Boxplots \newline
\rule[4pt]{\textwidth}{0.4pt}
```{r echo = FALSE}
rm(list = ls())
setwd('/Users/lschaack/Documents/cu/my_stuff/tips')
tipData = read.csv('tips.csv')
# clear dollar signs from dollar values to make them numeric
tipData$Tips = as.numeric(sub('$', '', as.character(tipData$Tips), fixed=TRUE))
tipData$Tips.per.hour = (tipData$Tips / tipData$Hours)
tipData$Index = seq(1, nrow(tipData))
# View(tipData)
# store valuable subsets
amTipData = tipData[tipData$AM.PM == 'AM', ]
pmTipData = tipData[tipData$AM.PM == 'PM', ]
assBarTipData = tipData[tipData$AM.PM == 'assBar', ]
midTipData = tipData[tipData$AM.PM == 'Mid', ]
# plots
par(mfrow = c(2, 2))
boxplot(x = tipData$Tips, ylim = c(20, 70), horizontal = TRUE, xlab = 'Tips in Dollars', main = 'All Tips')
boxplot(x = amTipData$Tips, ylim = c(20, 70), horizontal = TRUE, xlab = 'Tips in Dollars', main = 'AM Cashier Tips')
boxplot(x = assBarTipData$Tips, ylim = c(20, 70), horizontal = TRUE, xlab = 'Tips in Dollars', main = 'AssBar Tips')
boxplot(x = pmTipData$Tips, ylim = c(20, 70), horizontal = TRUE, xlab = 'Tips in Dollars', main = 'PM Cashier Tips')
```

## Averages \newline
\rule[4pt]{\textwidth}{0.4pt}
```{r echo = FALSE}
# Averages
univTipAvg = mean(tipData$Tips)
amTipAvg = mean(amTipData$Tips)
pmTipAvg = mean(pmTipData$Tips)
assBarTipAvg = mean(assBarTipData$Tips)
midTipAvg = mean(midTipData$Tips)
# Per Hour Averages
univTipPHAvg = mean(tipData$Tips.per.hour)
amTipPHAvg = mean(amTipData$Tips.per.hour)
pmTipPHAvg = mean(pmTipData$Tips.per.hour)
assBarTipPHAvg = mean(assBarTipData$Tips.per.hour)
midTipPHAvg = mean(midTipData$Tips.per.hour)
# beta distribution here? since values are so sparse...
# show distribution of shifts?
```
\begin{center}
\begin{tabular}{lccl}
\textbf{Shift} & \textbf{Total} & \textbf{Per Hour} & \\
\hline \\
Overall & \$`r signif(univTipAvg, digits = 4)` & \$`r signif(univTipPHAvg, digits = 3)` & `r length(tipData$Tips)` observations \\
AM Cashier & \$`r signif(amTipAvg, digits = 4)` & \$`r signif(amTipPHAvg, digits = 3)` & `r length(tipData[tipData$AM.PM == 'AM', ]$Tips)` observations \\
PM Cashier & \$`r signif(pmTipAvg, digits = 4)` & \$`r signif(pmTipPHAvg, digits = 3)` & `r length(tipData[tipData$AM.PM == 'PM', ]$Tips)` observations \\
Assistant Bar & \$`r signif(assBarTipAvg, digits = 4)` & \$`r signif(assBarTipPHAvg, digits = 3)` & `r length(tipData[tipData$AM.PM == 'assBar', ]$Tips)` observations \\
Mid (Baseline) & \$`r signif(midTipAvg, digits = 4)` & \$`r signif(midTipPHAvg, digits = 3)` & `r length(tipData[tipData$AM.PM == 'Mid', ]$Tips)` observations \\
\end{tabular}
\end{center}

```{r echo = FALSE}
weekdays = c('Sunday', 'Saturday', 'Friday', 'Thursday', 'Wednesday', 'Tuesday', 'Monday')
getAvg = function(weekday) {
	return(mean(tipData[tipData$Day.of.Week == weekday, ]$Tips))
}
byDay = lapply(weekdays, getAvg)
barplot(t(byDay), names.arg = substring(weekdays, first = 1, last = 3),
		  horiz = TRUE, las = 1, col = 'azure3', xlab = 'Average Dollars Per Hour',
		  main = 'Tips by Day of the Week')
```
\newline
\vfill
The Tuesday tip rate is inordinately high due to one of only a few observations coinciding with Thanksgiving.
\vfill
```{r echo = FALSE}
counts = factor(tipData$Day.of.Week)
counts = factor(counts, levels(counts)[c(2, 6, 7, 5, 1, 3, 4)])
plot(counts, names.arg = substring(rev(weekdays), first = 1, last = 3), xlab = 'Day of the Week', ylab = 'Number of Shifts Worked', main = 'Number of Shifts Worked By Day')
```

## Over Time (haw) \newline
\rule[4pt]{\textwidth}{0.4pt}
```{r echo = FALSE}
{
	plot(x = tipData$Index, y = tipData$Tips, ylim = c(0, 100), xlab = "Shifts Worked", ylab = "Tips")
	lines(x = amTipData$Index, y = amTipData$Tips, col = 'red')
	lines(x = pmTipData$Index, y = pmTipData$Tips, col = 'blue')
	lines(x = assBarTipData$Index, y = assBarTipData$Tips, col = 'green')
	lines(x = midTipData$Index, y = midTipData$Tips, col = 'orange')
	legend(x = "topleft", legend = c("AM Cashier", "PM Cashier", "Assistant Bar", "Mid"), fill = c("red", "blue", "green", "orange"))
}
```
\newline
\vfill
```{r echo = FALSE}
{
	plot(x = tipData$Index, y = tipData$Tips.per.hour, ylim = c(0, 15), xlab = "Shifts Worked", ylab = "Tips Per Hour")
	lines(x = amTipData$Index, y = amTipData$Tips.per.hour, col = 'red')
	lines(x = pmTipData$Index, y = pmTipData$Tips.per.hour, col = 'blue')
	lines(x = assBarTipData$Index, y = assBarTipData$Tips.per.hour, col = 'green')
	lines(x = midTipData$Index, y = midTipData$Tips.per.hour, col = 'orange')
	legend(x = "topleft", legend = c("AM Cashier", "PM Cashier", "Assistant Bar", "Mid"), fill = c("red", "blue", "green", "orange"))
}
```

## Multiple linear regression \newline
\rule[4pt]{\textwidth}{0.4pt}
```{r echo = FALSE}
# Potential to be improved by transforming shift name into a "start time" column...
# ...which, partnered with the Hours feature, might show nicer behavior.
lmod = lm(Tips ~ Hours + factor(AM.PM) + factor(Day.of.Week), data=tipData)
summary(lmod)
```

\vfill
```{r echo = FALSE}
{
	plot(x = tipData$Index, y=tipData$Tips, ylim = c(0, 100),
		  col = ifelse(tipData$AM.PM == 'AM', 'red',
		  		  ifelse(tipData$AM.PM == 'PM', 'blue',
		  		  ifelse(tipData$AM.PM == 'assBar', 'green',
		  		  ifelse(tipData$AM.PM == 'Mid', 'orange',
		  		  		 'yellow')))), # holiday
		  xlab = "Shifts Worked", ylab = "Tips")
	lines(tipData$Index, predict(lmod))
	legend(x = "topleft", legend = c("AM Cashier", "PM Cashier", "Assistant Bar", "Mid", "Holiday"), fill = c("red", "blue", "green", "orange", "yellow"))
}
```

## Prediction for next shift \newline
\rule[4pt]{\textwidth}{0.4pt}
```{r}
next.shift = data.frame(AM.PM = c('PM'), Day.of.Week = c('Saturday'), Hours = c(7.0))
prediction = predict(lmod, newdata = next.shift)
```
The model, such as it is, predicts \$`r signif(prediction, digits = 4)` $\;\pm\;$ \$`r signif(sigma(lmod), digits = 3)` in tips for the coming shift.
